What We Eval in the Shadows

### 2021-08-10 paper/main.tex

\paragraph{Discussion}

It is interesting how many different forms any given site sees. The more forms,
the harder it is to characterize behavior at that site. Luckily,
\packageNbOneMinimizedPercent of sites only see a single form. In other words,
a compiler could correctly predict the \emph{shape} of an expression by looking
just at the first call in \packageNbOneMinimizedPercent of the sites. Few sites
are highly polymorphic (8 or more different forms). These include the pipe
operator of the \c{magrittr} package, used to compose functions, \eg instead of
\c{f(g(h(x),y),z)} one can write \c{h(x) \%>\% g(y) \%>\% f(z)}.

The variety of uses of \eval is evidenced by the number of different forms. In
comparison, JavaScript \eval usage was more straightforward and more
predictable as reported by \citet{oopsla12b}, with 98.7\% of the sites with
only one form. On the other hand, there are \packageNbSimpleMinimizedOne sites
that only receive one of the simple forms (\ie\xspace \c{X}, \c{V}, \c{\$} or
\c{<-}). Additional \packageNbSimpleMinimizedMore sites takes multiple forms,
but all are these simple ones. Among these sites there are many cases where
\eval could be replaced by less powerful constructs such as \c{get} to lookup a
name in an environment, \c{assign} to set a value to a name in an environment
or \c{do.call} to reflectively call a function.

@@

\paragraph{Discussion}
The origin data suggests that reflection is the first use case for \eval, \ie using \c{match.call}, \c{substitute}, or \c{match.call} to access and transform the arguments of a function when calling the function. Constructing expressions from strings is also
a common use case. The constructed category shows that most
\evals comes from code that was processed by the compiler, and may be slightly
modified by the programmer before invoking \eval. Both constructed and
reflection categories roughly correspond to meta-programming. Some of these use
cases could likely be replaced by macros if the R designers could be convinced
to overcome their distaste for those. Another use of \eval is to directly build symbols to evaluate them in another environment than the default one.

>>>>

\mypara{Discussion.} 
It is interesting how many different forms any given site sees. The more forms,
the harder it is to characterize behavior at that site. Luckily,
\packageNbOneMinimizedPercent of sites only see a single form. In other words,
a compiler could correctly predict the \emph{shape} of an expression by looking
just at the first call in \packageNbOneMinimizedPercent of the sites. Few sites
are highly polymorphic (8 or more different forms). These include the pipe
operator of the \c{magrittr} package, used to compose functions, \eg instead of
\c{f(g(h(x),y),z)} one can write \c{h(x) \%>\% g(y) \%>\% f(z)}.

The number of different forms evidences the variety of uses of \eval. In
comparison, JavaScript \eval usage was more straightforward and more
predictable as reported by \citet{oopsla12b}, with 98.7\% of the sites with
only one form. On the other hand, there are \packageNbSimpleMinimizedOne sites
that only receive one of the simple forms (\ie\xspace \c{X}, \c{V}, \c{\$} or
\c{<-}). Additional \packageNbSimpleMinimizedMore sites take multiple forms,
but all of them are these simple ones. Among these sites, there are many cases
where \eval could be replaced by simpler constructs such as \c{get} to lookup a
name in an environment, \c{assign} to set a value to a name in an environment,
or \c{do.call} to call a function reflectively.

@@

\mypara{Discussion}
The origin data suggests that reflection is the first use case for \eval, \ie using \c{match.call}, \c{substitute}, or \c{match.call} to access and transform the arguments of a function when calling the function. Constructing expressions from strings is also
a common use case. The constructed category shows that most
\evals comes from code that was processed by the compiler, and may be slightly
modified by the programmer before invoking \eval. Both constructed and
reflection categories roughly correspond to meta-programming. Some of these use
cases could likely be replaced by macros if the R designers could be convinced
to overcome their distaste for those. Another use of \eval is to directly build symbols to evaluate them in another environment than the default one.
